---
title: "wine"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

# import library
```{r}
library(tidyverse)
library(moderndive)
library(ggplot2)
library(ggrepel)
library(grid)
library(gridExtra)
library(GGally)
library(skimr)
library(kableExtra)
library(ggpubr)
library(corrplot)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(forcats)
library(scales)
library(janitor)
```

|---------------------------------------------------------------------------------------------------------------|
| I. DATA EXTRACTION - Dataset cleansing                                                                        |
|---------------------------------------------------------------------------------------------------------------|

```{r}
# load the dataset
wine = read.csv("D:/COURSE/STATS5085 DATA ANALYSIS SKILLS/Assessment-Presentation/dataset18.csv", header = T)
wine = wine %>%
  select(-X)
summary(wine)
glimpse(wine)
```
```{r}
# omit observations with NA value & add new variable grade (categorize points whether >=90 or < 90)
wine_noNA <- wine %>% 
  mutate(grade = ifelse(wine$points>=90,'greater than 90','less than 90')) %>% 
  drop_na()
summary(wine_noNA)
glimpse(wine_noNA)
```
Comment : 
1. There are 7 variables in the wine dataset, which consist of 2 numerical variables and 5 categorical variables.
2. Points variable is the response and the rest are the explanatory variables.
3. The response variable need to be classified as the binary response into 2 categories (>=90 or <90).
4. The dataset contains 2000 observations, but there are only 1,844 observations after omitting the NA value.
5. The levels in each categorical variable are too large (26 countries, 141 provinces, 178 varieties, 1712 wineries, and 1997 titles), this implies that it is impossible taking all the levels into account in the analysis.

|---------------------------------------------------------------------------------------------------------------|
| I. DATA EXTRACTION - Levels Simplification                                                                    |
|---------------------------------------------------------------------------------------------------------------|

```{r}
# create country frequency
country_freq = wine_noNA %>%
  select(country) %>%
  group_by(country) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  mutate(relative_frequency = percent(frequency/sum(frequency)))
head(country_freq, 10)
view(country_freq)

# create province frequency
province_freq = wine_noNA %>%
  select(province) %>%
  group_by(province) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  mutate(relative_frequency = percent(frequency/sum(frequency)))
head(province_freq, 10)
view(province_freq)

# create variety frequency
variety_freq = wine_noNA %>%
  select(variety) %>%
  group_by(variety) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  mutate(relative_frequency = percent(frequency/sum(frequency))) 
head(variety_freq, 10)
view(variety_freq)

# create winery frequency
winery_freq = wine_noNA %>%
  select(winery) %>%
  group_by(winery) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  mutate(relative_frequency = percent(frequency/sum(frequency)))
head(winery_freq, 10)
view(winery_freq)

# create title frequency
title_freq = wine_noNA %>%
  select(title) %>%
  group_by(title) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  mutate(relative_frequency = percent(frequency/sum(frequency)))
head(title_freq, 10)
view(title_freq)
```

Comment : 
1. Using a specific threshold from the relative frequency (proportion), we can simplify the levels in each categorical variable.
2. After looking at the relative frequency distribution over the categorical variables, we decide to use the value of 4% as the threshold.
3. Using a 4% threshold means that each level with a relative frequency under 4% will be considered the new category called "The Other".
4. New Country Level: US, France, Italy, Other.
5. New Province Level: California, Washington, Tuscany, Oregon, Other.
6. New Variety Level : Pinot Noir, Chardonnay, Cabernet Sauvignon, Red Blend, Sauvignon Blanc, Bordeaux-style Red Blend, other.
7. There is no new categorical level for winery and title since all the levels of the relative frequency in these variables are under 4% (even 1%), so these variables will be treated just as identity observations and don't taken into account in the analysis.

```{r}
# set new level in each variable
wine_newlevel = wine_noNA %>%
  mutate(country = recode(country, "US"="US", "France"="France", "Italy"="Italy", .default="Other")) %>%
  mutate(province = recode(province, "California"="California", "Washington"="Washington", "Oregon"="Oregon", "Tuscany"="Tuscany", .default="Other")) %>%
  mutate(variety = recode(variety, "Pinot Noir"="Pinot Noir", "Chardonnay"="Chardonnay", "Cabernet Sauvignon"="Cabernet Sauvignon", "Red Blend"="Red Blend", "Sauvignon Blanc"="Sauvignon Blanc", "Bordeaux-style Red Blend"="Bordeaux-style Red Blend", .default="Other")) %>%
  select(grade, price, country, province, variety)

# set level order to assign other as baseline category
wine_newlevel$country = factor(wine_newlevel$country, levels = c("Other", "US", "France", "Italy"))
wine_newlevel$province = factor(wine_newlevel$province, levels = c("Other", "California", "Washington", "Oregon", "Tuscany"))
wine_newlevel$variety = factor(wine_newlevel$variety, levels = c("Other", "Pinot Noir", "Chardonnay", "Cabernet Sauvignon", "Red Blend", "Sauvignon Blanc", "Bordeaux-style Red Blend"))
wine_newlevel$grade = factor(wine_newlevel$grade, levels = c("less than 90", "greater than 90"))

head(wine_newlevel,10)
summary(wine_newlevel)
```

|---------------------------------------------------------------------------------------------------------------|
| II. DATA EXPLORATION - Statistic Summary                                                                      |
|---------------------------------------------------------------------------------------------------------------|

```{r}
# create statistics price
my_skim = skim_with(base = sfl(n = length))
stat_price = wine_newlevel %>%
  select(grade, price) %>%
  group_by(grade) %>%
  my_skim() %>%
  transmute(Grade=c('Less than 90', 'Greater than 90'), n=n,
            Mean=numeric.mean,
            St.Dev=numeric.sd, Min=numeric.p0, Q1=numeric.p25, Median=numeric.p50, Q2=numeric.p75, Max=numeric.p100) %>%
  kable(booktabs = TRUE, linesep = "", digits = 2) %>%
  kable_styling(font_size = 12, full_width = FALSE) %>%
  row_spec(0, color = "white", background = "navy") %>%
  row_spec(1:2, color = "black", background = "lightyellow", bold=TRUE)

# create contingency table of country
cont_country = wine_newlevel %>%
  group_by(country, grade) %>%
  summarize(frequency=n()) %>%
  spread(grade, frequency) %>%
  kable(booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size = 12, full_width = FALSE) %>%
  row_spec(0, color = "white", background = "navy") %>%
  row_spec(1:4, color = "black", background = "lightyellow", bold=TRUE)

# create contingency table of province
cont_province = wine_newlevel %>%
  group_by(province, grade) %>%
  summarize(frequency=n()) %>%
  spread(grade, frequency) %>%
  kable(booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size = 12, full_width = FALSE) %>%
  row_spec(0, color = "white", background = "navy") %>%
  row_spec(1:5, color = "black", background = "lightyellow", bold=TRUE)

# create contingency table of variety
cont_variety = wine_newlevel %>%
  group_by(variety, grade) %>%
  summarize(frequency=n()) %>%
  spread(grade, frequency) %>%
  kable(booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size = 12, full_width = FALSE) %>%
  row_spec(0, color = "white", background = "navy") %>%
  row_spec(1:7, color = "black", background = "lightyellow", bold=TRUE)

stat_price
cont_country
cont_province
cont_variety
```

|---------------------------------------------------------------------------------------------------------------|
| II. DATA EXPLORATION - Data Visualization                                                                     |
|---------------------------------------------------------------------------------------------------------------|

```{r}
# create boxplot price vs grade
boxplot_price = ggplot(wine_newlevel,aes(x=grade,y=price,fill=grade))+
  geom_boxplot() +
  coord_flip() +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'),
  legend.position = c(0.8, 0.2)) +
  scale_fill_manual(values = c("yellow2", "navy")) +
  labs(x="Grade", y="Price")

# subset data for creating barplot by country
wine_country = wine_newlevel %>%
  select(grade, country) %>%
  group_by(country, grade) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  mutate(perc = round(frequency/sum(frequency)*100, 2))
# create barplot by country
barplot_country = ggplot(wine_country, aes(x = reorder(country, frequency), y = frequency, fill = grade)) +
  geom_col() +
  coord_flip() +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'),
  legend.position = c(0.8, 0.2)) +
  scale_fill_manual(values = c("yellow2", "navy")) +
  labs(x = "Country", y = "Frequency", title = "Wine frequency by country")

# subset data for creating barplot by province
wine_province = wine_newlevel %>%
  select(grade, province) %>%
  group_by(province, grade) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  mutate(perc = round(frequency/sum(frequency)*100, 2))
# create barplot by province
barplot_province = ggplot(wine_province, aes(x = reorder(province, frequency), y = frequency, fill = grade)) +
  geom_col() +
  coord_flip() +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'),
  legend.position = c(0.8, 0.2)) +
  scale_fill_manual(values = c("yellow2", "navy")) +
  labs(x = "Province", y = "Frequency", title = "Wine frequency by province")

# subset data for creating barplot by variety
wine_variety = wine_newlevel %>%
  select(grade, variety) %>%
  group_by(variety, grade) %>%
  summarise(frequency = n()) %>%
  arrange(desc(frequency)) %>%
  mutate(perc = round(frequency/sum(frequency)*100, 2))
# create barplot by variety
barplot_variety = ggplot(wine_variety, aes(x = reorder(variety, frequency), y = frequency, fill = grade)) +
  geom_col() +
  coord_flip() +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'),
  legend.position = c(0.8, 0.2)) +
  scale_fill_manual(values = c("yellow2", "navy")) +
  labs(x = "Variety", y = "Frequency", title = "Wine frequency by variety")

boxplot_price
barplot_country
barplot_province
barplot_variety
```
Comment:
1. It can be seen that the price of  wine in the group with a points greater than or equal to 90 is often higher than that in the group with a points less than 90.
2. At the same time, there are many high priced wines in the "1" group, while the prices of wines in the "0" group are always less than 150 pounds. It can be seen that good wine is usually not cheap, and wine with low score is generally not too expensive.
3. The “Other” category has the most frequency in each barplot, except wine frequency by country.
4. Overall, the “less than 90” group has a higher number than “greater than 90” in each level category.


|---------------------------------------------------------------------------------------------------------------|
| III. MODEL SELECTION - Fitting single model                                                                   |
|---------------------------------------------------------------------------------------------------------------|

# Grade ~ Price
```{r}
# fit the first model
model_price1 = glm(grade ~ price, family = binomial, data=wine_newlevel)
summ(model_price1, digits = 3)

# obtain 95% confidence interval for the last iteration
confint(model_price1)

# plot points estimate and 95% confidence interval (Odds scale) for the last iteration
p1_price = plot_model(model_price1, show.values = TRUE, axis.lim = c(1,1.5), title="Odds (Price)", show.p = FALSE) +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'))

# plot probability of being awarded grade of greater than 90 for the last iteration
wine_price = wine_noNA %>% 
  mutate(price.probs=fitted(model_price1))
p2_price = ggplot(data=wine_price, aes(x=price, y=price.probs))+
  geom_smooth(method="glm",method.args=list(family="binomial"),se=FALSE)+
  labs(x="price",y="probability of wine being awarded grade of >= 90") +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'))

# laying out multiple plots on a page 
grid.arrange(p1_price, p2_price, nrow=1)
```
comment :
1. all terms are significant at the first iteration (grade ~ intercept + price).
2. the logodds of the "grade" increased by 0.058 ,if the price increases by one unit.
3. The odds is 1.06, which has a corresponding 95% confidence interval of (1.053,1.067) so for every 1 unit increase  in price, the odds of "grade" has grade>=90 is increased by 1.06.


# Grade ~ Country
```{r}
# fit first model
model_country1 = glm(grade ~ country, family = binomial, data=wine_newlevel)
summ(model_country1)

# set Italy to other category
wine_country2 = wine_newlevel %>%
  mutate(country = recode(country,
                             "US" = "US", "France" = "France", .default = "Other"))
# refit the second model
model_country2 = glm(grade ~ country, family = binomial, data=wine_country2)
summ(model_country2)

# obtain 95% confidence interval for the last iteration
confint(model_country2)

# plot points estimate and 95% confidence interval (Odds scale)
p1_country = plot_model(model_country2, show.values = TRUE, axis.lim = c(1,1.5), title="Odds (Country)", show.p = FALSE) +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'))

# plot Prob. of wine being awarded grade of >= 90
p2_country = plot_model(model_country2, type = "pred", title = "", axis.title = c("Country", "Prob. of wine being awarded grade of >= 90"))

p1_country
p2_country
```
comment:
All terms are significant at the second iteration (grade ~ intercept + countryUS + countryFrance)

# Grade ~ Province
```{r}
# fit first model
model_province1 = glm(grade ~ province, family = binomial, data=wine_newlevel)
summ(model_province1)

# set Oregon & Tuscany to other category
wine_province2 = wine_newlevel %>%
  mutate(province = recode(province,
                             "California" = "California", "Washington" = "Washington", .default = "Other"))
# refit the second model
model_province2 = glm(grade ~ province, family = binomial, data=wine_province2)
summ(model_province2)

# obtain 95% confidence interval for the last iteration
confint(model_province2)

# plot points estimate and 95% confidence interval (Odds scale)
p1_province = plot_model(model_province2, show.values = TRUE, axis.lim = c(1,1.5), title="Odds (Province)", show.p = FALSE) +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'))

# plot Prob. of wine being awarded grade of >= 90
p2_province = plot_model(model_province2, type = "pred", title = "", axis.title = c("Province", "Prob. of wine being awarded grade of >= 90"))

p1_province
p2_province
```
comment:
All terms are significant at the second iteration (grade ~ intercept + provinceCalifornia + provinceWashington)

# Grade ~ Variety
```{r}
# fit first model
model_variety1 = glm(grade ~ variety, family = binomial, data=wine_newlevel)
summ(model_variety1)

# set Chardonnay, Cabernet Sauvignon, Red Blend, Sauvignon Blanc, and Bordeaux-style Red Blend to other category
wine_variety2 = wine_newlevel %>%
  mutate(variety = recode(variety,
                             "Pinot Noir" = "Pinot Noir", "Bordeaux-style Red Blend" = "Bordeaux-style Red Blend", .default = "Other"))
# refit the second model
model_variety2 = glm(grade ~ variety, family = binomial, data=wine_variety2)
summ(model_variety2)

# obtain 95% confidence interval for the last iteration
confint(model_variety2)

# plot points estimate and 95% confidence interval (odds scale)
p1_variety = plot_model(model_province2, show.values = TRUE, axis.lim = c(1,1.5), title="Odds(Variety)", show.p = FALSE) +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'))

# plot Prob. of wine being awarded grade of >= 90
p2_variety = plot_model(model_variety2, type = "pred", title = "", axis.title = c("Variety", "Prob. of wine being awarded grade of >= 90"))

p1_variety
p2_variety
```
comment:
All terms are significant at the second iteration (grade = intercept + varietyPinot Noir + varietyBordeaux-style Red Blend)


|---------------------------------------------------------------------------------------------------------------|
| III. MODEL SELECTION - Fitting combination model                                                              |
|---------------------------------------------------------------------------------------------------------------|

# Grade ~ all
```{r}
# fit first model
model_combination1 = glm(grade ~ ., family = binomial, data=wine_newlevel)
summ(model_combination1)

# set France, Oregon, Tuscany, Pinot Noir, Chardonnay, Red Blend, Sauvignon Blanc, and Bordeaux-style Red Blend to other category
wine_combination2 = wine_newlevel %>%
  mutate(country = recode(country,
                                "US" = "US", "Italy" = "Italy", .default = "Other")) %>%
  mutate(province = recode(province,
                                "California" = "California", "Washington" = "Washington", .default = "Other")) %>%
  mutate(variety = recode(variety,
                             "Cabernet Sauvignon" = "Cabernet Sauvignon", .default = "Other"))
# refit the second model
model_combination2 = glm(grade ~ ., family = binomial, data=wine_combination2)
summ(model_combination2)

# set California to other category
wine_combination3 = wine_newlevel %>%
  mutate(country = recode(country,
                          "US" = "US", "Italy" = "Italy", .default = "Other")) %>%
  mutate(province = recode(province,
                          "Washington" = "Washington", .default = "Other")) %>%
  mutate(variety = recode(variety,
                          "Cabernet Sauvignon" = "Cabernet Sauvignon", .default = "Other"))
# refit the third model
model_combination3 = glm(grade ~ ., family = binomial, data=wine_combination3)
summ(model_combination3)

# set US to other category
wine_combination4 = wine_newlevel %>%
  mutate(country = recode(country,
                          "Italy" = "Italy", .default = "Other")) %>%
  mutate(province = recode(province,
                          "Washington" = "Washington", .default = "Other")) %>%
  mutate(variety = recode(variety,
                          "Cabernet Sauvignon" = "Cabernet Sauvignon", .default = "Other"))
# refit the forth model
model_combination4 = glm(grade ~ ., family = binomial, data=wine_combination4)
summ(model_combination4)

# obtain 95% confidence interval for the last iteration
confint(model_combination4)

# plot points estimate and 95% confidence interval (odds ratio)
p1_combination = plot_model(model_combination4, show.values = TRUE, axis.lim = c(0.2,5),title="Odds (Combination)", show.p = FALSE) +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'))

wine_combination_price = wine_noNA %>% 
  mutate(price.probs=fitted(model_combination4))
p2_combination = ggplot(data=wine_combination_price, aes(x=price, y=price.probs))+
  geom_smooth(method="glm",method.args=list(family="binomial"),se=FALSE)+
  labs(x="price",y="probability of wine being awarded grade of >= 90") +
  theme(panel.background = element_rect(fill = 'white', color = 'grey', size=0.5),
  panel.grid.major = element_line(color = 'grey', linetype = 'dotted'))

# plot Prob. of wine being awarded grade of >= 90
p3_combination = plot_model(model_combination4, type = "pred", title = "", axis.title = c("Categorical", "Prob. of wine being awarded grade of >= 90"))

p1_combination
p2_combination
p3_combination
```
comment:
all terms are significant at the forth iteration (grade ~ intercept + price + countryItaly + provinceWashington + varietyCabernet Sauvignon)


|---------------------------------------------------------------------------------------------------------------|
| III. MODEL SELECTION - AIC & BIC Comparison                                                                   |
|---------------------------------------------------------------------------------------------------------------|

```{r}
broom::glance(model_price1)
broom::glance(model_country2)
broom::glance(model_province2)
broom::glance(model_variety2)
broom::glance(model_combination4)
```

